Forecasting biodiversity in breeding birds using best practices

### 2017-12-07 manuscript/reviewer_reply.Rmd

## Editors Comments

*The reviewers and I agree that this manuscript is interesting, timely and well-written. I will add a few of my own observations here. First, and least important - normally submitted manuscripts do not hyphenate and do not right justify.*

**Thank you. We have removed the right justification and hyphenation.**

@@

*Line 107: For the environmental variables, do you have any that match up with the temporal resolution of the response data (monthly?). It seems like all the variables are annual / quarterly / etc.*

**See comment above on using shorter time scale environmental variables.**

*Line 124: I’m not as familiar with terrestrial projections, but are the CMIP5 projections relatively unbiased / precise when compared to recent data?*

**The accuracy of CMIP5 models varies by the variable considered. Temperature is represented very well over North America. Precipitation is challenging, especially over mountainous terrain, but the CMIP5 ensemble represents precipitation across North America relatively well ("Evaluation of precipitation and temperature simulation performance of the CMIP3 and CMIP5 historical experiments", Koutroulis et al. 2016).**

### 2017-12-07 manuscript/reviewer_reply.Rmd

*Table 1: Just my personal preference, but I’d include the check marks and replace the ‘Xs’ with blank cells*

**We have made this change**

*Line 194-221: It seems for each approach you only considered the full models with all environmental covariates? I’m curious whether any of the covariates appeared to not improve explanatory power, and whether you considered models that dropped these uninformative predictors?*

**We considered dropping uninformative predictors, but decided not to because of the complexity it would add to our analyses. In particular, it wasn't clear how to assess variable importance in a consistent way across three different model types and 500 Monte Carlo samples, and we did not want to bias the results by choosing a variable selection method that worked better for one method than another.**

### 2017-12-07 manuscript/reviewer_reply.Rmd

*Line 62: I really like the recommendations / clarity in Box 1, and think these will be accessible to most ecologists*

>Thank you.

*Line 71-73: I’m re-reading this after commenting on data, and am now a little confused. The bird responses are annual snapshots during the breeding season? And I thought the environmental data you used was annual or quarterly, not monthly?* https://www.mbr-pwrc.usgs.gov/bbs/grass/genintro.htm

>BBS data are generally viewed as annual data, even though the observations happen on a single day. The presence of an individual species is primarily a function of suitable habitat and, for many species, timing of migration. Environmental variables with a shorter temporal scale could likely improve models, such as incorporating cool weather days into the observation model, and would be worth pursuing in a different study.

@@

*Line 92: For those not familiar with the BBS protocol, maybe describe in a sentence. I could be mis-remembering, but isn’t each site made up of 50 stops, and presumably you aggregated richness across these stops?*

>This is correct, we've added clarification to the Methods.

*Line 105: I assume all these species have been similarly ID’d over the course of the dataset – e.g. birds split into sub-species haven’t been included.*

>This is correct, see the section "Richness Data" in the Methods.

@@

*Line 107: For the environmental variables, do you have any that match up with the temporal resolution of the response data (monthly?). It seems like all the variables are annual / quarterly / etc.*

>See comment above on using shorter time scale environmental variables.

@@

>We have made this change

*Line 194-221: It seems for each approach you only considered the full models with all environmental covariates? I’m curious whether any of the covariates appeared to not improve explanatory power, and whether you considered models that dropped these uninformative predictors?*

*Line 216: Cite randomForest / caret/ party R package – or whatever else you used?*

>The randomForest package was cited in the final methods paragraph. We've cited it where it is first mentioned as well.

@@

*Here are some comments:*

*Birds are not the best system for doing any sort of distribution analyses. They are incredibly vagile, hard to detect, and the distributions of breeding birds are affected by many components besides environmental variables (particularly interspecific interactions). No assessment of detectability is included here other than accounting for observer identity. However, I do realize that BBS data are detailed and compelling data to use given what a large amount of detailed time-series it provides.*

*Abstract: The authors forecast the data and then test it using known data. Hindcast only refers to back-in-time predictions.*

>The definition of hindcasting varies by field. We use it here as it's used in meteorology, where any prediction of observations already available is generally a hindcast, and forecasts are only for predictions which are actually in the future. See Jolliffe & Stephenson 2003 (in manuscript) section 1.4.2 for a longer discussion. We've clarified this in the text.

@@

#### Methods:

*the authors perform their analyses on data that are collected from an incredibly dynamic time period. I wouldn’t expect species to have settled into a richness pattern given the incredibly disrupted nature of the last 32 years. Some strategies to try to handle these issues might be to train and predict at different temporal scales and to both forecast and hindcast the predictions. (e.g. train on 2004-2013 and try to predict 1982-1992 data + 1993-2003 data). If the authors were to make splits based on expected periods of dynamism, they could better tell whether the drivers that are included in the models are accurate predictors.*

*124-130: I didn’t see where future projections were performed/used in this paper. Why is this here?*

>Thank you for pointing out that we were unclear here.  The methods and purpose for the future predictions are now explained in more detail on lines [[XX]].

### 2017-12-07 manuscript/reviewer_reply.Rmd

#### Introduction:
*Line 41: I think this is generally more common in terrestrial ecology – in fisheries these SDM models with more of a theoretical background don’t seem to be used as much*

*Line 62: I really like the recommendations / clarity in Box 1, and think these will be accessible to most ecologists*

*Line 71-73: I’m re-reading this after commenting on data, and am now a little confused. The bird responses are annual snapshots during the breeding season? And I thought the environmental data you used was annual or quarterly, not monthly?* https://www.mbr-pwrc.usgs.gov/bbs/grass/genintro.htm

**BBS data are generally viewed as annual data, even though the observations happen on a single day. The presence of an individual species is primarily a function of suitable habitat and, for many species, timing of migration. Environmental variables with a shorter temporal scale could likely improve models, such as incorporating cool weather days into the observation model, and would be worth pursuing in a different study.**

@@

*Line 138: Maybe also cite any of the interfaces you used, like RStan: http://mc-stan.org/users/citations/*

**RStan is cited in the last paragraph of the Methods.**

*Line 140: I’m curious why you didn’t consider spatial structure in the random effects here, using GMRF models or similar? One hand these models might be overkill if the habitat/environmental variables (which are spatially correlated) explain variability in richness, but the spatial correlation may pick up other sources of variability not accounted for by those variables.*

>The goal of this model is to control for observer effects, not to provide a focal model of the system (though we do use it for the average model as a convenience). Because observers typically observe the same site year after year it was necessary to include site in this model to avoid classifying variation among sites as differences among observers and incorrectly removing it before fitting other models. Given the use of this model, it is unclear what including spatial structure would accomplish for the broader goals of the paper. Moreover, it seems possible that it could end up correcting out meaningful ecological patterns and therefore confuse the comparison of more complex models to the baselines.

@@

** We have made this change**

*Line 194-221: It seems for each approach you only considered the full models with all environmental covariates? I’m curious whether any of the covariates appeared to not improve explanatory power, and whether you considered models that dropped these uninformative predictors?*

*Line 216: Cite randomForest / caret/ party R package – or whatever else you used?*

**The randomForest package was cited in the final methods paragraph. We've cited it where it is first mentioned as well.**

@@

*Here are some comments:*

*Birds are not the best system for doing any sort of distribution analyses. They are incredibly vagile, hard to detect, and the distributions of breeding birds are affected by many components besides environmental variables (particularly interspecific interactions). No assessment of detectability is included here other than accounting for observer identity. However, I do realize that BBS data are detailed and compelling data to use given what a large amount of detailed time-series it provides.*

*Abstract: The authors forecast the data and then test it using known data. Hindcast only refers to back-in-time predictions.*

**The definition of hindcasting varies by field. We use it here as it's used in meteorology, where any prediction of observations already available is generally a hindcast, and forecasts are only for predictions which are actually in the future. See Jolliffe & Stephenson 2003 (in manuscript) section 1.4.2 for a longer discussion. We've clarified this in the text.**

### 2017-12-08 manuscript/reviewer_reply.Rmd

*In general, I thought this paper was interesting and well written. These results add to the increasing literature supporting the idea that ecological forecasting in time/space is difficult. I think the recommendations/best practice bullet points described by the authors are essential for anyone interested in forecasting this kind of data. All of my points below can be considered minor issues, which I think will improve the paper’s clarity.*

>Thank you.

#### Introduction:
*Line 41: I think this is generally more common in terrestrial ecology – in fisheries these SDM models with more of a theoretical background don’t seem to be used as much*

>This is a good point. We have added the word "terrestrial" to this line.

@@

*Line 107: For the environmental variables, do you have any that match up with the temporal resolution of the response data (monthly?). It seems like all the variables are annual / quarterly / etc.*

>See comment above on using shorter time scale environmental variables.

*Line 124: I’m not as familiar with terrestrial projections, but are the CMIP5 projections relatively unbiased / precise when compared to recent data?*

>The accuracy of CMIP5 models varies by the variable considered. Temperature is represented very well over North America. Precipitation is challenging, especially over mountainous terrain, but the CMIP5 ensemble represents precipitation across North America relatively well ("Evaluation of precipitation and temperature simulation performance of the CMIP3 and CMIP5 historical experiments", Koutroulis et al. 2016).

@@

>RStan is cited in the last paragraph of the Methods.

*Line 140: I’m curious why you didn’t consider spatial structure in the random effects here, using GMRF models or similar? One hand these models might be overkill if the habitat/environmental variables (which are spatially correlated) explain variability in richness, but the spatial correlation may pick up other sources of variability not accounted for by those variables.*

*Line 141-143: I think the paper thus far is pretty clear. I was hoping to see a few equations here – it helps me especially when components start to become hierarchical!*

>We have added the equations and Stan code as a supplemental file.

@@

*Table 1: Just my personal preference, but I’d include the check marks and replace the ‘Xs’ with blank cells*

>We have made this change

*Line 194-221: It seems for each approach you only considered the full models with all environmental covariates? I’m curious whether any of the covariates appeared to not improve explanatory power, and whether you considered models that dropped these uninformative predictors?*

>We considered dropping uninformative predictors, but decided not to because of the complexity it would add to our analyses. In particular, it wasn't clear how to assess variable importance in a consistent way across three different model types and 500 Monte Carlo samples, and we did not want to bias the results by choosing a variable selection method that worked better for one method than another.

@@

*Line 216: Cite randomForest / caret/ party R package – or whatever else you used?*

>The randomForest package was cited in the final methods paragraph. We've cited it where it is first mentioned as well.

*Line 242: Again, I think including the equation here would be useful*

>We have added an in-line equation and clarified the wording

@@

*124-130: I didn’t see where future projections were performed/used in this paper. Why is this here?*

>Thank you for pointing out that we were unclear here.  The methods and purpose for the future predictions are now explained in more detail on lines [[XX]].

*Fig. 2: Panels not labeled. No key for colors. Somewhat confusing.*

>Thank you for pointing this out.  As the Editor noted, we had inadvertently reversed the two panels. We have fixed this, and also labeled them as you suggested.

### 2017-12-08 manuscript/observer-equations.Rmd
auto-arima details in the suppliment

>>>>

**Details of the auto.arima model**

We used non-seasonal time series model $ARIMA(p,d,q)$, with drift allowed, in the function 'auto.arima' of the R package `forecast`. The search space for $p$ and $q$ were $(2,5)$ and $(2,5)$, respectively. $d$ was selected via KPSS unit-root tests as described in Hyndman and Khandakar (2008) and with a maximum value of $2$. Stepwise model selection was done using AICc as the model criteria.
